Evolutionary Design of Particles for Collectively Intelligent Problem Solving

### 2014-05-08 docs/paper.tex

The set of states in each controller was fixed to be the same set of compound behaviors as those found in \cite{rodriguez2004extending}, but with parameters derived through evolutionary programming. This allowed for easy comparison with the authoritative, human-derived solution. In addition, the transitions between states were kept the same except for the guarding thresholds, which we evolved. New individuals were introduced through the application of mutation and recombination operators.

The mutation operator first selected which parameters to modify independently. We used a 20\% chance for mutating each weight, radius, alpha, and threshold. Each selected threshold had an equal chance of increasing by 1 and decreasing by 1. Each selected real-valued parameter was modified through the addition of a uniformly distributed random number with mean 0. We chose to mutate each weight by -0.2 to 0.2 and each alpha and radius by -20 to 20.

The recombination operator also ignored the elite individuals carried over from the last population. Parents were selected with a likelihood of 0.2 to be recombined with another individual. Our recombination operator replaced the first parent with intermediary recombination, that is the mean value of the parents velocity parameters. The second chosen parent remained in the population. Although elite individuals were exempt from recombination, they were able to be selected as the secondary parent, contributing their genetic material to the recombined offspring.

A round of evolution combined the mutation and recombination operators to produce the next generation. After evaluating every individual in the current generation, we applied elitism, tournament selection, mutation, and recombination to complete a round of evolution. For elitism, the single best individual of the previous generation was always carried into the next generation without modification. The other 49 individuals were selected via tournaments of size 3. Each of the 49 was then subjected to the mutation operator, with its parameters being mutated independently.

>>>>

The set of states in each controller was fixed to be the same set of compound behaviors as those found in \cite{rodriguez2004extending}, but with parameters derived through evolutionary programming. This allowed for easy comparison with the authoritative, human-derived solution. In addition, the transitions between states were kept the same except for the guarding thresholds, which were evolved. New individuals were introduced through the application of mutation and recombination operators.

The mutation operator first selected which parameters to modify independently. We used a 20\% chance for mutating each weight, radius, alpha, and threshold. Each selected threshold had an equal chance of increasing by 1 and decreasing by 1. Each selected real-valued parameter was modified through the addition of a uniformly distributed random number with mean 0. We mutated each weight by -0.2 to 0.2 and each alpha and radius by -20 to 20.

The recombination operator also ignored the elite individuals carried over from the last population. Parents were selected with a likelihood of 0.2 to be recombined with another individual. We used intermediary recombination of the weight, radius, and alpha parameters of all components in all states, bounded by the minimum and maximum possible values of each parameter. The second chosen parent remained in the population. Although elite individuals were exempt from recombination, they were able to be selected as the secondary parent, contributing their genetic material to the recombined offspring.

A round of evolution consisted of the evaluation of every individual followed by the application of elitism, tournament selection, recombination, and mutation to produce the next generation. For elitism, the single best individual of the previous generation was always carried into the next generation without modification. The other 49 individuals were selected via tournaments of size 3. The 49 individuals chosen through the tournament then each had a 20\% chance of undergoing recombination with another randomly chosen child. Each of the resulting 49 was then subjected to the mutation operator, mutating parameters independently.
